---
output: pdf_document
classoption: portrait
---
\pagenumbering{gobble}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dr)
library(dplyr)
library(ggplot2)
library(ggbeeswarm)
library(grid)
library(gridExtra)
```


```{r data}
load(file = 'ex14_results.rda')
```

The bias from IPW (i.e. `inferference`), outcome, and doubly robust estimators of Liu et al was assessed for the following data generating scenarios. $n_i = 30$ for $m = 300$ groups. The simulation was run on 300 datasets.

Treatment was simulated from:
\[
A_{ij} \sim Bern(\mbox{logit}^{-1}(\gamma_0 + \gamma_1 |Z_{1ij}| + \gamma_2 Z_{2ij} + \gamma_3 |Z_{1ij}| Z_{2ij} + b_i))
\]

where $Z_{1ij} \sim N(0, 1)$, $Z_{2ij} \sim Bern(0.5)$, and $b_i \sim N(0, \theta^2)$.

The outcome was simulated from:

\[
Y_{ij} \sim N(\beta_0 + \beta_1 A_{ij} + \beta_2 f(A_i) + \beta_3 Z_{1ij} + \beta_4 Z_{2ij} + \beta_5 |Z_{1ij}| Z_{2ij}, 1)
\]

where $f(A_i)$ is the proportion of subjects treated in group $i$.

```{r}
knitr::kable(scenarios %>% select(-m, -ni), caption = 'Parameter scenarios',
             col.names = c('Scenario', '$\\gamma$', '$\\theta$', '$\\beta$'))
```

\newpage

```{r, fig.height=6, fig.cap="Showing mean bias for the simulations on the -log10 scale, so to the right is less bias, left is more. The dotted line corresponds with 0.01. Vertical facets are scenarios. Horizontal facets are target estimands."}
results %>%
  filter(!is.na(a), model_spec == 'tcor_ocor') %>%
  mutate(target = paste0('Y(', a, ',', alpha, ')'),
         val    = paste0(round(mean_bias, 3), ' (', round(ese, 3), ')'),
         logbias = -log10(abs(mean_bias))) %>%
  ggplot(
    data = .,
    aes(x = -log10(abs(mean_bias)),
        y = method, 
        shape = method,
        color = method)
  ) + 
  geom_vline(xintercept = 2, linetype = 'dotted') + 
  geom_point() + 
  scale_y_discrete(name = '') + 
  scale_x_continuous(name = expression('mean(log'[10]*'(|bias|))')) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  scale_shape(guide = FALSE) + 
  facet_grid(sid ~ target) + 
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7)
  )
  # ungroup() %>%
  # select(sid, method, target, val) %>%
  # tidyr::spread(target, value = val) %>%
  # knitr::kable()
```

\newpage

```{r make_plot_fun}
# Make plot to examine bias for a scenario
make_bias_plot <- function(.sid){
  plot_dt <- estimates %>% filter(sid == .sid, model_spec == 'tcor_ocor', !is.na(a)) %>%
    mutate(method_label = paste(method, hajek))
  
  failures <- plot_dt %>% group_by(method_label, alpha, a) %>% 
    summarise(failed = sum(failed)) %>%
    mutate(fail_label = ifelse(failed == 0, '', as.character(failed)))
  
  ggplot(
    data = plot_dt,
    aes(x = factor(alpha), y = log10(abs(bias)))) +
    # geom_boxplot(outlier.size = 0.5) +
    geom_violin(scale = 'width') + 
    geom_text(
      data = failures,
      aes(x = factor(alpha), y = -6, label = fail_label),
      size = 2,
      inherit.aes = FALSE
    ) +
    scale_y_continuous(
      name = expression('log'[10]*'(|bias|)'),
      breaks = c(-6, -4, -2, 0, 2),
      labels = c('Failures', -4, -2, 0, 2)
    ) +
    coord_cartesian(
      ylim = c(-6, 3)
    ) + 
    scale_x_discrete(
      name = expression(alpha)
    ) + 
  facet_grid(a ~ method_label)
}

# Make plots to demonstrate example of scenario
make_example_plots <- function(...){
  dots <- list(...)
  args <- dots[pmatch(names(dots), names(formals((gen_data))))]
  sampledt    <- do.call(gen_data, args)
  
  p1 <- ggplot(data = sampledt, aes(x = p)) + 
    geom_histogram(bins = 30 ) + 
    scale_x_continuous(limits = c(0, 1)) +
    ggtitle('Distribution of Pr(A|Z)') + 
    theme(
      plot.title = element_text(size = 9)
    )
  
  p2 <- ggplot(data = distinct(sampledt, group, fA), aes(x = fA)) + 
    geom_histogram(bins = 20) + 
    scale_x_continuous(limits = c(0, 1)) + 
    ggtitle('Distribution of f(A)') + 
    theme(
      plot.title = element_text(size = 9)
    )
 arrangeGrob(
  grobs = list(ggplotGrob(p1), ggplotGrob(p2)), 
  ncol = 1,
  widths = unit(2.5, 'in'))
}

# Arrange all plots for a scenario
make_scenario_plots <- function(.sid){
  dargs <- filter(scenarios, sid == .sid)
  pp <-make_example_plots(
    m = 300, 
    ni = 20,
    gamma = dargs$gamma[[1]], 
    theta = dargs$theta[[1]], 
    beta  = dargs$beta[[1]])
  
  row2 <- arrangeGrob(
      grobs = list(
        pp, 
        ggplotGrob(make_bias_plot(.sid))),
      widths = unit(c(2.25, 4.5), 'in'),
      ncol = 2,
      padding = unit(0, 'lines'))
  
  # Organize final output
  arrangeGrob(
    grobs = list(
      textGrob(label = paste('Scenario', .sid),
               x = 0,
               just = 'left'),
      row2),
      nrow = 2,
      heights = unit(c(.5, 4), 'in')
  )
}
```

```{r}
plots <- make_scenario_plots(1)
grid.draw(plots)
```
