---
output: pdf_document
classoption: portrait
---
\pagenumbering{gobble}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
load(file = 'ex11_results.rda')
library(dplyr)
library(ggplot2)
library(ggbeeswarm)
library(grid)
library(gridExtra)
```

The bias from IPW (i.e. `inferference`), outcome, and doubly robust estimators of Liu et al was assessed for the following data generating scenarios. $n_i = 20$ for $m = 300$ groups. The simulation was run on 250 datasets..

Treatment was simulated from:
\[
A_{ij} \sim Bern(\mbox{logit}^{-1}(\gamma_0 + \gamma_1 Z_{1ij} + \gamma_2 Z_{2ij} + b_i))
\]

where $Z_{1ij} \sim N(0, 1)$, $Z_{2ij} \sim Bern(0.5)$, and $b_i \sim N(0, \theta^2)$.

The outcome was simulated from:

\[
Y_{ij} \sim N(\beta_0 + \beta_1 A_{ij} + \beta_2 f(A_i) + \beta_3 Z_{1ij} + \beta_4 Z_{2ij}, 1)
\]

where $f(A_i)$ is the proportion of subjects treated in group $i$.

```{r}
knitr::kable(scenarios, caption = 'Parameter scenarios',
             col.names = c('Scenario', '$\\gamma$', '$\\theta$', '$\\beta$'))
```

\newpage

```{r, fig.height=6, fig.cap="Showing mean bias for 250 simulations on the -log10 scale, so to the right is less bias, left is more. The dotted line corresponds with 0.01. Vertical facets are scenarios. Horizontal facets are target estimands."}
results %>%
  mutate(target = paste0('Y(', a, ',', alpha, ')'),
         val    = paste0(round(mean_bias, 3), ' (', round(ese, 3), ')'),
         logbias = -log10(abs(mean_bias))) %>%
  ggplot(
    data = .,
    aes(x = -log10(abs(mean_bias)),
        y = method, 
        shape = method,
        color = method)
  ) + 
  geom_vline(xintercept = 2, linetype = 'dotted') + 
  geom_point() + 
  scale_y_discrete(name = '') + 
  scale_x_continuous(name = expression('log[10](|bias|)')) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  scale_shape(guide = FALSE) + 
  facet_grid(sid ~ target) + 
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7)
  )
  # ungroup() %>%
  # select(sid, method, target, val) %>%
  # tidyr::spread(target, value = val) %>%
  # knitr::kable()
```

\newpage

```{r gen_data_fun}
gen_data <- function(m, ni, gamma, theta, beta){
  n <- m * ni
  data_frame(
    group = rep(1:m, each = ni),
    Z1  = rnorm(n, sd = 1),
    Z2  = rbinom(n, size = 1, prob = .5),
    b   = rep(rnorm(m, sd = theta), each = ni),
    p   = as.numeric(plogis(cbind(1, Z1, Z2) %*% gamma + b)),
    A   = rbinom(n, size = 1, prob = p),
    fA  = rep(tapply(A, group, mean), each = ni),
    Y   = as.numeric(cbind(1, A, fA, Z1, Z2) %*% beta)
  )
}
```

```{r make_plot_fun}
# Make plot to examine bias for a scenario
make_bias_plot <- function(.sid){
  plot_dt <- estimates %>% filter(sid == .sid)
  failures <- plot_dt %>% group_by(method, alpha, a) %>% 
    summarise(failed = sum(failed)) %>%
    mutate(fail_label = ifelse(failed == 0, '', as.character(failed)))

  ggplot(
    data = plot_dt,
    aes(x = factor(alpha), y = log10(abs(bias)))) +
    # geom_boxplot(outlier.size = 0.5) +
    geom_violin(scale = 'width') + 
    # geom_text(
    #   data = failures,
    #   aes(x = factor(alpha), y = 2, label = fail_label),
    #   inherit.aes = FALSE
    # ) +
    scale_y_continuous(
      name = expression('log'[10]*'(|bias|)'),
      limits = c(-6, 3)
    ) +
    scale_x_discrete(
      name = expression(alpha)
    ) + 
  facet_grid(a ~ method)
}

# Make plots to demonstrate example of scenario
make_example_plots <- function(...){
  dots <- list(...)
  args <- dots[pmatch(names(dots), names(formals((gen_data))))]
  sampledt    <- do.call(gen_data, args)
  
  p1 <- ggplot(data = sampledt, aes(x = p)) + 
    geom_histogram(bins = 30 ) + 
    scale_x_continuous(limits = c(0, 1)) +
    ggtitle('Distribution of Pr(A|Z)') + 
    theme(
      plot.title = element_text(size = 9)
    )
  
  p2 <- ggplot(data = distinct(sampledt, group, fA), aes(x = fA)) + 
    geom_histogram(bins = 20) + 
    scale_x_continuous(limits = c(0, 1)) + 
    ggtitle('Distribution of f(A)') + 
    theme(
      plot.title = element_text(size = 9)
    )
 arrangeGrob(
  grobs = list(ggplotGrob(p1), ggplotGrob(p2)), ncol = 1)
}

# Arrange all plots for a scenario
make_scenario_plots <- function(.sid){
  dargs <- filter(scenarios, sid == .sid)
  pp <-make_example_plots(
    m = 300, 
    ni = 20,
    gamma = dargs$gamma[[1]], 
    theta = dargs$theta[[1]], 
    beta  = dargs$beta[[1]])
  
  row2 <- arrangeGrob(
      grobs = list(
        pp, 
        ggplotGrob(make_bias_plot(.sid))),
      widths = unit(c(2.5, 4.5), 'in'),
      ncol = 2)
  
  # Organize final output
  arrangeGrob(
    grobs = list(
      textGrob(label = paste('Scenario', .sid),
               x = 0,
               just = 'left'),
      row2),
      nrow = 2,
      heights = unit(c(.5, 4), 'in')
  )
}
```

```{r}
plots <- lapply(1:12, function(i) make_scenario_plots(i))
```


```{r}
grid.draw(plots[[1]])
```

```{r}
grid.draw(plots[[2]])
```

```{r}
grid.draw(plots[[3]])
```

```{r}
grid.draw(plots[[4]])
```

```{r}
grid.draw(plots[[5]])
```

```{r}
grid.draw(plots[[6]])
```

```{r}
grid.draw(plots[[7]])
```

```{r}
grid.draw(plots[[8]])
```

```{r}
grid.draw(plots[[9]])
```

```{r}
grid.draw(plots[[10]])
```

```{r}
grid.draw(plots[[11]])
```

```{r}
grid.draw(plots[[12]])
```







