---
title: "Coding DR estimator"
output: pdf_document
---

```{r filehead, echo = FALSE}
#   Title: Doubly Robust Estimator with Interference
#  Author: B. Saul
#    Date: 2016-03-01
# Purpose: code the DR estimator in Liu et al
```

```{r libraries, echo=TRUE, message=FALSE, warning=FALSE}
library(inferference)
library(dplyr)
library(geepack)
library(magrittr)
#library(plyr)
```

To test the Doubly Robust (DR) estimator, I will use the `vaccinesim` data in the package `inferference`. This data has 250 groups, based on the simulations in Perez-Heydrich (2014).

Let $p_{ij}= p_{ij}(A_i) = \sum_{k \neq j} A_{ik} /(n_i - 1)$ the proportion treated in group $i$, excluding the $j$th person.

```{r, echo=TRUE}
vaccinesim <- vaccinesim %>%
  mutate(id = row_number()) %>%
  group_by(group) %>%
  # compute p_ij for each subject
  mutate(pA = (sum(A) - A)/(n() - 1)) %>%
  ungroup()
```

For the outcome model:
\[
\mu_{ij} = logit^{-1}(\beta_0 + \beta_1 A_{ij} + \beta_2 p_{ij} + \beta_3 A_{ij} p_{ij} + \beta_4 X_{1ij} + \beta_5 X_{2ij})
\]

$\beta$ will be estimated using GEE using an independence working correlation.

```{r outcome_model, echo=TRUE}
m_outcome <- geepack::geeglm(y ~ A*pA + X1 + X2, data = vaccinesim,
                             id = id, family = binomial)
```

For the ipw model:
\[
h_{ij} = logit^{-1}(\gamma_1 + \gamma_2 X_{1ij} + \gamma_3 X_{ij} + b_i)
\]

\[
f(A_i | X_i, \gamma) = \int \prod_{j} h_{ij}^{A_{ij}} (1 - h_{ij} )^{(1 - A_{ij})} df_{b_i}
\]

$\gamma$ is estimated from logistic-normal random effects model.

```{r ipw_model, echo = TRUE}
alphas <- c(.2, .3, .4, .5, .6, .7)

ipw_pieces <- interference(formula = y | A | B ~ X1 + X2 + (1|group) | group,
                     data = as.data.frame(vaccinesim), 
                     allocations = alphas,
                     method = 'simple')
```


Let $W_{i} = \frac{\pi(A_i, \alpha_k)}{f(A_i| X_i; \hat{\gamma})}$

Compute the first term of DR estimator:
\[
[Y_{ij}(A_i) - \mu_{ij}(A_i, X_i, \hat{\beta}) ] W_i
\]

```{r term1, echo=TRUE}
## First term ####
dr_term1 <- function(data, alpha, ipw_part, outcome_model)
{
  # Get the weights from the interference object
  weights <- ipw_part$weights[ , as.character(alpha)]
  weights <- data.frame(group = as.integer(names(weights)), w = weights)
  
  data %>%   
    # Compute mu_ij for each subject
    mutate_(muA = ~predict(outcome_model, type = 'response')) %>%
    # Add group IPW weights to the data
    left_join(weights, by = 'group') %>%
    # Compute term 1
    mutate_(term1 = ~ (y - muA) * w )
}
```

Compute the second term of DR estimator:
\[
\sum_{a_i} \mu_{ij}(a_i, X_i, \hat{\beta}) \pi(a_i; \alpha_k)
\]

```{r term2, echo=TRUE}
## Second term ####
dr_term2 <- function(data, alpha, outcome_model)
{
  data %>%
    select(-pA) %>%
    # this needs to be done by group. plyr splits the data frame by group
    # then applies a function to each piece
    {plyr::dlply(., plyr::.(group), function(x){
      x %>%
        # Generate all possible sum(a_i) for each subject
        merge(expand.grid(sum_a = 0:(n_distinct(.$id) - 1)), all = T)  %>%
        group_by_(~id) %>%
        # Compute pi and p_ij for each sum(a_i)
        mutate_(pA = ~ sum_a/(n() - 1),
                pi = ~ choose(n() - 1, sum_a) * alpha^sum_a * (1 - alpha)^(n() - 1 - sum_a)) %>%
        ungroup() %>%
        # Compute mu_ij for each a_i per subject
        # For some reason predict() is throwing error with geeglm model
        #mutate_(mu_ij = ~predict(outcome_model, newdata = ., type = 'response')) %>%
        mutate_(mu =~ as.numeric(
          plogis(model.matrix(outcome_model$formula, .) %*% coef(outcome_model) ) ) ) %>%
        # Sum by individual to compute term2
        group_by_(~id) %>%
        summarize_(term2 = ~sum(mu * pi)) 
    })} %>%
  # Put the split data back together
  bind_rows()
}

```

Putting term 1 and term 2 together:

\[
\hat{Y}_{ij}(\alpha_k) = [Y_{ij}(A_i) - \mu_{ij}(A_i, X_i, \hat{\beta}) ] W_i + \sum_{a_i} \mu_{ij}(a_i, X_i, \hat{\beta}) \pi(a_i; \alpha_k)
\]

Then average by group:
\[
\hat{Y}_{i}(\alpha_k) = \sum_j \hat{Y}_{ij}(\alpha_k)/n_i
\]

Then average across groups:
\[
\hat{Y}(\alpha_k) = \sum_i \hat{Y}_{i}(\alpha_k)/m
\]

```{r DR, echo = TRUE}
## All together now ####
dr <- function(data, alpha, ipw_part, outcome_model)
{
  dr_term1(data, alpha, ipw_part, outcome_model) %>%
    left_join(dr_term2(data, alpha, outcome_model), by = 'id') %>%
    mutate_(Y_hat_ij = ~term1 + term2) %>%
    group_by_(~group) %>%
    summarise_(Y_hat_i = ~mean(Y_hat_ij)) %>%
    summarise_(Y_hat = ~mean(Y_hat_i) )
}
```
Compute $\hat{Y}_{DR}(\alpha)$ for 6 different $\alpha$ levels: .2, .3, .4, .5, .6, .7.

```{r DR_estimate, echo=TRUE}
lapply(alphas, function(alpha) dr(vaccinesim, alpha, ipw_pieces, m_outcome)) %>%
  bind_rows()
```

How do these compare to $\hat{Y}_{IPW}(\alpha)$?

```{r IPW_estimate, echo=TRUE}
ipw_pieces$estimates %>%
  filter(effect_type == 'outcome', marginal == TRUE) %>%
  select(alpha1, estimate)
```

Why does the DR estimator return negative values?

\begin{align*}
\frac{(Y_{ij}(A_i) - \mu_{ij}(A_i) )}{f} \pi(A_i) + \sum_{a_i} \mu_{ij}(a_i) \pi(a_i) \\
\frac{(Y_{ij}(A_i) - \mu_{ij}(A_i) )}{f} \pi(A_i) + \mu_{ij}(A_i) + \sum_{a_i \neq A_i} \mu_{ij}(a_i) \pi(a_i) \\
\frac{(Y_{ij}(A_i) - \mu_{ij}(A_i) )}{f} \pi(A_i) + \mu_{ij}(A_i)\pi(A_i) + \sum_{a_i \neq A_i} \mu_{ij}(a_i) \pi(a_i) \\
\left\{ \frac{(Y_{ij}(A_i) - \mu_{ij}(A_i) )}{f} + \mu_{ij}(A_i) \right \}\pi(A_i) + \sum_{a_i \neq A_i} \mu_{ij}(a_i) \pi(a_i) \\
\left\{ \frac{Y_{ij}(A_i) - \mu_{ij}(A_i) + f \mu_{ij}(A_i))}{f}  \right \}\pi(A_i) + \sum_{a_i \neq A_i} \mu_{ij}(a_i) \pi(a_i) \\
\left\{ \frac{Y_{ij}(A_i) + \mu_{ij}(A_i)(f - 1)}{f} \right \}\pi(A_i) + \sum_{a_i \neq A_i} \mu_{ij}(a_i) \pi(a_i) \\
\end{align*}

Since $f - 1 \leq 0$, $\mu_{ij}(A_i) \in (0, 1)$, and $Y_{ij} \in \{0, 1\}$,

\[
Y_{ij}(A_i) + \mu_{ij}(A_i)(f - 1) 
\begin{cases}
< 0 & \text{if } Y = 0 \\
> 0 & \text{if } Y = 1
\end{cases}
\]

