---
output: pdf_document
classoption: portrait
---
\pagenumbering{gobble}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(dr)
library(dplyr)
library(ggplot2)
library(ggbeeswarm)
library(grid)
library(gridExtra)
```


```{r data}
load(file = 'ex19_results.rda')
```

The bias from IPW (i.e. `inferference`), outcome, and doubly robust estimators of Liu et al was assessed for the following data generating scenarios. The simulation was run on `r nsims` datasets.

Treatment was simulated from:
\[
A_{ij} \sim Bern(\mbox{logit}^{-1}(\gamma_0 + \gamma_1 |Z_{1ij}| + \gamma_2 Z_{2ij} + \gamma_3 |Z_{1ij}| Z_{2ij} + b_i))
\]

where $Z_{1ij} \sim N(0, 1)$, $Z_{2ij} \sim Bern(0.5)$, and $b_i \sim N(0, \theta^2)$.

The outcome was simulated from:

\[
Y_{ij} \sim N(\beta_0 + \beta_1 A_{ij} + \beta_2 f(A_i) + \beta_3 Z_{1ij} + \beta_4 Z_{2ij} + \beta_5 |Z_{1ij}| Z_{2ij}, 1)
\]

where $f(A_i)$ is the proportion of subjects treated in group $i$.

```{r}
knitr::kable(scenarios, caption = 'Data generating scenarios',
             col.names = c('Scenario', '$\\gamma$', '$\\theta$', '$\\beta$', '$m$', '$n_i$'))
```

\newpage

```{r, fig.height=6, fig.cap="Showing mean bias for the simulations on the -log10 scale, so to the right is less bias, left is more. The dotted line corresponds with 0.01. Vertical facets are scenarios. Horizontal facets are target estimands."}
results %>%
  filter(!is.na(a)) %>%
  mutate(target = paste0('Y(', a, ',', alpha, ')'),
         val    = paste0(round(mean_bias, 3), ' (', round(ese, 3), ')'),
         logbias = -log10(abs(mean_bias)),
         method_label = paste(method, hajek)) %>%
  ggplot(
    data = .,
    aes(x = -log10(abs(mean_bias)),
        y = method_label, 
        shape = method_label,
        color = method_label)
  ) + 
  geom_vline(xintercept = 2, linetype = 'dotted') + 
  geom_point() + 
  scale_y_discrete(name = '') + 
  scale_x_continuous(name = expression('mean(log'[10]*'(|bias|))')) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  scale_shape(guide = FALSE) + 
  facet_grid(model_spec + sid ~ target) + 
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7)
  )
  # ungroup() %>%
  # select(sid, method, target, val) %>%
  # tidyr::spread(target, value = val) %>%
  # knitr::kable()
```
\newpage
```{r, fig.height=6, fig.cap=""}
results %>%
  filter(!is.na(a)) %>%
  mutate(target = paste0('Y(', a, ',', alpha, ')'),
         val    = paste0(round(mean_bias, 3), ' (', round(ese, 3), ')'),
         logbias = -log10(abs(mean_bias)),
         method_label = paste(method, hajek)) %>%
  ggplot(
    data = .,
    aes(x = mean_bias,
        y = coverage, 
        shape = method_label,
        color = method_label)
  ) + 
  geom_point() + 
  scale_y_continuous(name = 'Coverage') + 
  scale_x_continuous(name = expression('mean(bias)')) + 
  scale_color_brewer(palette = 'Set1', guide = FALSE) + 
  scale_shape() + 
  facet_grid(model_spec + sid ~ target) + 
  theme(
    axis.text.x = element_text(size = 7),
    axis.text.y = element_text(size = 7)
  )
  # ungroup() %>%
  # select(sid, method, target, val) %>%
  # tidyr::spread(target, value = val) %>%
  # knitr::kable()
```

\newpage

```{r make_plot_fun}
# Make plot to examine bias for a scenario
make_bias_plot <- function(.sid, .model_spec, .show_x_labels, .show_y_labels){
  plot_dt <- estimates %>% 
    filter(sid == .sid, model_spec == .model_spec, !is.na(a)) %>%
    mutate(method_label = paste(method, hajek))
  
  failures <- plot_dt %>% group_by(method_label, alpha, a) %>% 
    summarise(failed = sum(failed)) %>%
    mutate(fail_label = ifelse(failed == 0, '', as.character(failed)))
  
  p <- ggplot(
    data = plot_dt %>% filter(failed == FALSE),
    aes(x = factor(alpha), y = log10(abs(bias)))) +
    geom_hline(yintercept = -2) +
    geom_hline(yintercept = 0) + 
    # geom_boxplot(outlier.size = 0.5) +
    geom_violin(scale = 'width') + 
    geom_text(
      data = failures,
      aes(x = factor(alpha), y = -6, label = fail_label),
      size = 2,
      inherit.aes = FALSE
    ) +
    scale_y_continuous(
      name = ifelse(.show_y_labels, expression('log'[10]*'(|bias|)'), ' '),
      breaks = c(-6, -4, -2, 0, 2),
      labels = c('Failures', -4, -2, 0, 2)
    ) +
    coord_cartesian(
      ylim = c(-6, 3)
    ) + 
    scale_x_discrete(
      name = ifelse(.show_x_labels, expression(alpha), ' ')
    ) + 
  facet_grid(a ~ method_label)
  
  if(!.show_x_labels){
    p <- p +
      theme(
        axis.text.x  = element_blank(),
        axis.ticks.x = element_blank()
      )
  }
  
  p
}

# Make plots to demonstrate example of scenario
make_example_plots <- function(...){
  dots <- list(...)
  args <- dots[pmatch(names(dots), names(formals((gen_data))))]
  sampledt    <- do.call(gen_data, args)
  
  p1 <- ggplot(data = sampledt, aes(x = p)) + 
    geom_histogram(bins = 30 ) + 
    scale_x_continuous(limits = c(0, 1)) +
    ggtitle('Distribution of Pr(A|Z)') + 
    theme(
      plot.title = element_text(size = 9)
    )
  
  p2 <- ggplot(data = distinct(sampledt, group, fA), aes(x = fA)) + 
    geom_histogram(bins = 20) + 
    scale_x_continuous(limits = c(0, 1)) + 
    ggtitle('Distribution of f(A)') + 
    theme(
      plot.title = element_text(size = 9)
    )
 arrangeGrob(
  grobs = list(ggplotGrob(p1), ggplotGrob(p2)), 
  ncol = 2,
  heights = unit(1.5, 'in'),
  widths = unit(c(2.5, 2.5), 'in'))
}

arrange_bias_plots <- function(.sid, row_height){
  
  nrows <- length(margs)
  # Make the bias plots for all model specifications
    bias_plots <- lapply(seq_along(margs), function(i) {
      spec <- names(margs[i])
      show_x <- ifelse(i == nrows, TRUE, FALSE)
      show_y <- ifelse(i == 1, TRUE, FALSE)
      p <- make_bias_plot(
        .sid = .sid, 
        .model_spec = spec,
        .show_x_labels = show_x,
        .show_y_labels = show_y) 
        
        arrangeGrob(
          grobs = list(
            ggplotGrob(p),
            textGrob(label = spec, rot = 270)),
          nrow = 1,
          ncol =2,
          widths  = unit(c(6, .2), 'in'),
          heights = unit(c(row_height), 'in')
        )
      })
  
  arrangeGrob(
    grobs   = bias_plots,
    nrow    = nrows,
    heights = unit(rep(row_height, nrows), 'in'),
    padding = unit(0, 'lines')
  )
}


# Arrange all plots for a scenario
make_scenario_plots <- function(.sid){
  dargs <- filter(scenarios, sid == .sid)
  pp <-make_example_plots(
    m = 300, 
    ni = 20,
    gamma = dargs$gamma[[1]], 
    theta = dargs$theta[[1]], 
    beta  = dargs$beta[[1]])
  
  all_plots <- arrangeGrob(
      grobs = list(
        pp, 
        arrange_bias_plots(.sid, 2.2)),
      # widths = unit(c(2.25, 4.5), 'in'),
      # ncol = 2,
      nrow = 2,
      heights = unit(c(1.5, 2.2*4), 'in'),
      padding = unit(0, 'lines'))
  
  # Organize final output
  arrangeGrob(
    grobs = list(
      textGrob(label = paste('Scenario', .sid),
               x = 0,
               just = 'left'),
      all_plots),
      nrow = 2,
      heights = unit(c(.5, 8), 'in')
  )
}
```

```{r, fig.height = 10.5}
plots <- make_scenario_plots(5)
grid.draw(plots)
```

```{r}

plot_one <- function(.model_spec, .method, .axis_labels = FALSE, .method_labels = FALSE){
  
  estimates <- estimates %>%
      ungroup() %>%
    filter_(~ model_spec == .model_spec, ~method == .method, ~alpha == 0.5, ~a == 1) %>%
    mutate(log10absbias = -log10(abs(bias)))

  summary_vals <- results %>% ungroup() %>%
    filter_(~ model_spec == .model_spec, ~method == .method, ~alpha == 0.5, ~a == 1) %>%
    mutate(log10absbias = -log10(abs(mean_bias))) 

  # Plot prep  
  y_coverage <- round(summary_vals$coverage, 2)
  
  if(.axis_labels){
    yaxis_text <- element_text(size = 7, hjust = 1)
  } else {
    yaxis_text <- element_blank()
  }
  
  if(.method == 'ipw'){
    color <- "#658b83"
  } else if(.method == 'otc'){
    color <- "#a4044d" 
  } else if(.method == 'dbr'){
    color <- "#359721"
  }
         
  ## Bias plot
  bias_plot <- ggplot(estimates, 
         aes(x = 1, y =log10absbias )) +
    geom_hline(yintercept = 2, linetype = 'dashed') +
    geom_hline(yintercept = 0, linetype = 'dotted' ) + 
    geom_violin(
      scale = 'width',
      fill = color
    ) +
    geom_point(data = summary_vals) + 
    scale_x_continuous(
      name = '',
      expand = c(0,0)
    ) + 
    scale_y_continuous(
      name   = '',
      breaks = -c(1, 0, -1, -2, -3),
      labels = c(10, 1, 0.1, 0.01, 0.001),
      expand = c(0, 0)
    ) + 
    coord_cartesian(ylim = -c(2, -6)) + 
    annotate(geom = 'text', label = '0.01 bias') + 
    theme_void() + 
    theme(
      plot.title   = element_blank(),
      panel.margin = unit(c(0, 0, 0, 0), 'cm'),
      plot.margin  = unit(c(0, 0, 0, 0), 'cm'),
      axis.text.y  = yaxis_text,
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x  = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    )

  if(.method_labels){
    bias_plot <- bias_plot +
      geom_text(
        data = summary_vals,
        aes(x = 1, y = -1, label = method),
                color = color)
  }
  
  coverage_plot <- ggplot(
    summary_vals,
    aes(x = 1, y = coverage)) +
    geom_bar(
      aes(x = 1, y = 1),
      stat = 'identity', width = .1,
      fill = NA,
      color = 'black',
      inherit.aes = FALSE
    ) + 
    geom_bar(
      stat = 'identity', 
      width = .1,
      fill = color
    ) +
    # geom_text(
    #   aes(label = round(coverage, 2), y = coverage + 0.01),
    #   size = 3
    # ) + 
    scale_x_continuous(
      name = '',
      expand = c(0, 0)
    ) + 
    scale_y_continuous(
      name ='',
      limits = c(0, 1),
      breaks = y_coverage,
      labels = y_coverage,
      expand = c(0, 0)
    ) + 
    theme_void() + 
    theme(
      plot.margin  = unit(c(0, 0, 0, 0), 'cm'),
      axis.text.y  = yaxis_text,
      axis.ticks.y = element_blank(),
      axis.title.y = element_blank(),
      axis.text.x  = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    )


  p3 <- arrangeGrob(
    ggplotGrob(bias_plot),
    ggplotGrob(coverage_plot),
    ncol = 1,
    widths = unit(c(.5), 'in'),
    heights = unit(c(2, 1), 'in'),
    padding = unit(0, 'lines'))
  
  p3

}

# plot_one('tcor_ocor', 'ipw', TRUE) %>% grid.draw()
# ggplot(
#   test2, 
#   aes(y = coverage)) +
#   geom_col()
```

```{r}
plot_two <- function(.model_spec, .method_labels){
  arrangeGrob(
  grobs = list(
    plot_one(.model_spec, 'ipw', FALSE, .method_labels),
    plot_one(.model_spec, 'otc', FALSE, .method_labels),
    plot_one(.model_spec, 'dbr', FALSE, .method_labels)),
  widths = unit(rep(.5, 3), 'in'),
  ncol = 3,
  padding = unit(0, 'lines'))
}

plot_three <- function(.model_spec, .spec_label, .method_labels){
  plots <- arrangeGrob(
  grobs = list(
    textGrob(label = .spec_label, 
             just = 'left',
             x = 0),
    plot_two(.model_spec, .method_labels)),
  nrow = 2,
  widths = unit(1.5, 'in'),
  heights = unit(c(.2, 3), 'in'))
  
  rect <- rectGrob(
    # height = unit(c(0.15, row_height*rows), 'in'), 
    # width  = unit(8/3 * cols, 'in'), 
    height = unit(1, 'npc'), 
    width  = unit(1, 'npc'), 
    gp = gpar(lwd = 1, col = "grey80", fill = NA)) 
  
  gTree(children = gList(plots, rect))
}


```

```{r}
bias_plot_axes <- ggplot(
    data_frame(log10absbias = c(10, .000001)), 
         aes(x = 1, y =log10absbias )) +
    geom_blank() + 
    scale_x_continuous(
      name = '',
      expand = c(0,0)
    ) + 
    scale_y_continuous(
      name   = expression('-log'[10]*'(|bias|)'),
      breaks = -c(1, 0, -1, -2, -3),
      labels = c(10, 1, 0.1, 0.01, 0.001),
      expand = c(0, 0)
    ) + 
    coord_cartesian(ylim = -c(2, -6)) + 
    theme_void() + 
    theme(
      plot.title   = element_blank(),
      panel.margin = unit(c(0, 0, 0, 0), 'cm'),
      plot.margin  = unit(c(0, 0, 0, 0), 'cm'),
      axis.text.y  = element_text(size = 7, hjust = 1),
      axis.ticks.y = element_blank(),
      axis.title.y = element_text(size = 7, hjust = 1),
      axis.text.x  = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    )

coverage_plot_axes <- ggplot(
    data_frame(coverage = c(0, 1)),
    aes(x = 1, y = coverage)) +
    geom_blank() + 
    scale_x_continuous(
      name = '',
      expand = c(0, 0)
    ) + 
    scale_y_continuous(
      name ='Coverage',
      limits = c(0, 1),
      breaks = c(0, 1),
      expand = c(0, 0)
    ) + 
    theme_void() + 
    theme(
      plot.margin  = unit(c(0, 0, 0, 1), 'cm'),
      axis.text.y  = element_text(size = 7, hjust = 1),
      axis.ticks.y = element_blank(),
      axis.title.y = element_text(size = 7, hjust = 1),
      axis.text.x  = element_blank(),
      axis.ticks.x = element_blank(),
      axis.title.x = element_blank()
    )
axes <- arrangeGrob(
    ggplotGrob(bias_plot_axes),
    ggplotGrob(coverage_plot_axes),
    ncol = 1,
    widths = unit(c(.75), 'in'),
    heights = unit(c(2, 1), 'in'),
    padding = unit(0, 'lines'))

axes_aligned <- arrangeGrob(
  grobs = list(
    textGrob(label = '', 
             just = 'left',
             x = 0),
    axes),
  nrow = 2,
  widths = unit(.75, 'in'),
  heights = unit(c(.2, 3), 'in'))

```

```{r}
grid.newpage()


plot_four <- function(){
  p1 <- plot_three('tcor_ocor', 'T: correct\nO: correct', .method_labels = TRUE)
  p2 <- plot_three('tmis_ocor', 'T: misspecified\nO: correct', .method_labels = FALSE)
  p3 <- plot_three('tcor_omis', 'T: correct\nO: misspecified', .method_labels = FALSE)
  p4 <- plot_three('tmis_omis', 'T: misspecified\nO: misspecified', .method_labels = FALSE)
  arrangeGrob(
  axes_aligned, p1, p2, p3, p4,
  widths = unit(c(.8, rep(.5*3 + .25, 4)), 'in'),
  ncol = 5)
}

arrangeGrob(
  textGrob(label = expression('Y(1, 0.5)')),
  plot_four(),
  nrow = 2,
  heights = unit(c(.5, 3.5), 'in')
) %>% grid.draw()


```
